---
title: "p8105_hw6_sdc2157"
author: "Stephanie Calluori"
date: 2023-12-02
output: github_document
---

# Load packages and set seed
```{r packages, message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(readr)

set.seed(1)

```

```{r setup, message = FALSE, echo = FALSE, results = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

closed by arrest (solved) is 1; unsolved (wihtout arrest, open) 0; wanted to make numeric and binary

```{r}
homicide_raw <- read_csv("data/homicide-data.csv", col_names = TRUE, na = c("", "NA", "Unknown"))

homicide_clean <- homicide_raw |> 
  janitor::clean_names() |> 
  mutate(victim_age = as.numeric(victim_age)) |> 
  mutate(city_state = paste(city, state, sep = ", ")) |> 
  filter(!city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")) |> 
  filter(victim_race == "White" | victim_race == "Black") |> 
  mutate(
    victim_race = fct_relevel(victim_race, "White"),
    victim_sex = fct_relevel(victim_sex, "Female"),
    resolved = as.numeric(disposition == "Closed by arrest"))

```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
baltimore_df <- homicide_clean |> 
  filter(city_state == "Baltimore, MD")

fit_baltimore <- glm(resolved ~ victim_age + victim_sex + victim_race, data = baltimore_df, family = binomial())

fit_baltimore |> 
  broom::tidy(conf.int = TRUE) |> 
  rename(log_OR = estimate, log_conf_low = conf.low, log_conf_high = conf.high) |> 
  mutate(OR = exp(log_OR),
         conf_low = exp(log_conf_low),
         conf_high = exp(log_conf_high)) |> 
  select(term, OR, conf_low, conf_high, p.value) |> 
  knitr::kable(digits = 3)
  

```
In Baltimore, cases in which the victim is male have a 0.426 lower odds of being solved compared to cases in which the victim is female.
95% CI (0.324, 0.558)

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
homicide_nest_df <- homicide_clean |> 
  select(city_state, everything()) |> 
  nest(.data = _, data = uid:resolved)

results <- homicide_nest_df |> 
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())),
    results = map(models, \(mod) broom::tidy(x = mod, conf.int = TRUE))
  ) |> 
  unnest(results) |> 
  rename(log_OR = estimate, log_conf_low = conf.low, log_conf_high = conf.high) |> 
  mutate(OR = exp(log_OR),
         conf_low = exp(log_conf_low),
         conf_high = exp(log_conf_high)) |> 
  select(city_state, term, OR, conf_low, conf_high) |> 
  filter(term == "victim_sexMale") |> 
  arrange(desc(OR))

```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
results |> 
  mutate(city_state = forcats::fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(
    aes(x = city_state,
        ymin = conf_low,
        ymax = conf_high)
  ) +
  labs(
    title = "Adjusted OR for solving homicides comparing Male to Female victims in each city",
    x = "city_state",
    y = "estimated OR"
  ) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1))

```

# Problem 2
## Load Central Park Weather Data
```{r}
cp_weather_df <-
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

strap (ie our bootstrap sample)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words

```{r}
cp_results <- cp_weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_1 = map(models, broom::tidy),
    results_2 = map(models, broom::glance)
    ) |> 
  unnest(results_1) |> 
  select(term, estimate, results_2) |> 
  unnest(results_2) |> 
  select(term, estimate, r.squared)
  
cp_results_tidy <- cp_results |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(log_coefficients_mult = log(tmin * prcp))

```
for each model, we have an r.squared value, an intercept, tmin, prcp, and log_coefficients_mult

r^2 distribution
```{r}
cp_results_tidy |> 
  ggplot(aes(x = r.squared)) +
  geom_density()

```

r^2 distribution = partially left skewed. centers around 0.92. about 92% of the variation in tmax can be explained by the linear relationship between tmax and tmin + prcp

```{r}
cp_results_tidy |> 
  filter(log_coefficients_mult == "NaN") |> 
  nrow()

cp_results_tidy |> 
  filter(!log_coefficients_mult == "NaN") |> 
  ggplot(aes(x = log_coefficients_mult)) +
  geom_density()

```
noted total number of NAs produced; filtered them out for the distribution

log of product of coefficients distribution = centers around -5.5; What does this mean?

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2
 and log(β̂ 0∗β̂ 1)
. 
```{r}
cp_results_tidy |> 
  select(r.squared, log_coefficients_mult) |> 
  pivot_longer(
    r.squared:log_coefficients_mult,
    names_to = "term",
    values_to = "estimate"
  ) |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025, na.rm = TRUE),
    ci_upper = quantile(estimate, 0.975, na.rm = TRUE))
```


# Problem 3
there are no actual values of unknown for frace and no actual values of other for mrace so removed those from the fct levels
no NA values present
birthweight is a continuous outcome so use lm instead of glm

going with model 4

HOW TO USE ADD PREDICTIONS???
```{r}
birthweight_raw <- read_csv("data/birthweight.csv", col_names = TRUE)

birthweight_clean <- birthweight_raw |> 
  mutate(babysex = recode(
    babysex,
    "1" = "male",
    "2" = "female"),
    frace = recode(
      frace,
      "1" = "White",
      "2" = "Black",
      "3" = "Asian",
      "4" = "Puerto Rican",
      "8" = "Other",
      "9" = "Unknown"),
    mrace = recode(
      mrace,
      "1" = "White",
      "2" = "Black",
      "3" = "Asian",
      "4" = "Puerto Rican",
      "8" = "Other"),
    malform = recode(
      malform,
      "0" = "absent",
      "1" = "present")
    ) |> 
  mutate(
    babysex = forcats::fct_relevel(babysex, c("male", "female")),
    frace = forcats::fct_relevel(frace, c("White", "Black", "Asian", "Puerto Rican", "Other")),
    mrace = forcats::fct_relevel(mrace, c("White", "Black", "Asian", "Puerto Rican")),
    malform = forcats::fct_relevel(malform, c("absent", "present"))
  )

sum(is.na(birthweight_clean))

str(birthweight_clean)

fit_1 = lm(bwt ~ gaweeks, data = birthweight_clean)
fit_1 |> 
  broom::tidy()

birthweight_clean |> 
  ggplot(aes(x = gaweeks, y = bwt)) + 
  geom_point()

birthweight_clean |> 
  modelr::add_residuals(fit_1) |> 
  ggplot(aes(x = smoken, y = resid)) + geom_violin()


fit_1 = lm(bwt ~ gaweeks, data = birthweight_clean)
fit_2 = lm(bwt ~ gaweeks + bhead, data = birthweight_clean)
fit_3 = lm(bwt ~ gaweeks + bhead + blength, data = birthweight_clean)

fit_4 = lm(bwt ~ gaweeks + bhead + blength + babysex, data = birthweight_clean)
fit_5 = lm(bwt ~ gaweeks + bhead + blength + malform, data = birthweight_clean)
fit_6 = lm(bwt ~ gaweeks + bhead + blength + smoken, data = birthweight_clean)

anova(fit_1, fit_2) |> 
  broom::tidy()

anova(fit_2, fit_3) |> 
  broom::tidy()

anova(fit_3, fit_4) |> 
  broom::tidy()

anova(fit_3, fit_5) |> 
  broom::tidy()

anova(fit_3, fit_6) |> 
  broom::tidy()

birthweight_clean |> 
  modelr::add_residuals(fit_4) |> 
  ggplot(aes(x = gaweeks, y = resid)) + geom_violin()

birthweight_clean |> 
  modelr::add_residuals(fit_4) |> 
  ggplot(aes(x = bhead, y = resid)) + geom_violin()

birthweight_clean |> 
  modelr::add_residuals(fit_4) |> 
  ggplot(aes(x = blength, y = resid)) + geom_violin()

birthweight_clean |> 
  modelr::add_residuals(fit_4) |> 
  ggplot(aes(x = babysex, y = resid)) + geom_violin()


fit_4 = lm(bwt ~ gaweeks + bhead + blength + babysex, data = birthweight_clean)
fit_alt1 = lm(bwt ~ gaweeks + blength, data = birthweight_clean)
fit_alt2 = lm(bwt ~ blength * bhead * babysex, data = birthweight_clean)

fit_alt2 |> 
  broom::tidy()

birthweight_clean |> 
  modelr::add_residuals(fit_alt1) |> 
  ggplot(aes(x = gaweeks, y = resid)) + geom_violin()

birthweight_clean |> 
  modelr::add_residuals(fit_alt1) |> 
  ggplot(aes(x = blength, y = resid)) + geom_violin()

cv_df <-
  crossv_mc(birthweight_clean, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df <-
  cv_df |> 
  mutate(
    mod_alt1  = map(train, \(df) lm(bwt ~ gaweeks + bhead + blength + babysex, data = df)),
    mod_alt2  = map(train, \(df) lm(bwt ~ gaweeks + blength, data = df)),
    mod_alt3  = map(train, \(df) lm(bwt ~ blength*bhead*babysex, data = as_tibble(df)))) |> 
  mutate(
    rmse_alt1 = map2_dbl(mod_alt1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_alt2 = map2_dbl(mod_alt2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_alt3 = map2_dbl(mod_alt3, test, \(mod, df) rmse(model = mod, data = df)))


cv_df |> 
  select(rmse_alt1, rmse_alt2, rmse_alt3) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()


```







