---
title: "p8105_hw6_sdc2157"
author: "Stephanie Calluori"
date: 2023-12-02
output: github_document
---

# Load packages and set seed
```{r packages, message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)

```

```{r setup, message = FALSE, echo = FALSE, results = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1



# Problem 2
## Load Central Park Weather Data
```{r}
cp_weather_df <-
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

strap (ie our bootstrap sample)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words

```{r}
cp_results <- cp_weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_1 = map(models, broom::tidy),
    results_2 = map(models, broom::glance)
    ) |> 
  unnest(results_1) |> 
  select(term, estimate, results_2) |> 
  unnest(results_2) |> 
  select(term, estimate, r.squared)
  
cp_results_tidy <- cp_results |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(log_coefficients_mult = log(tmin * prcp))

```
for each model, we have an r.squared value, an intercept, tmin, prcp, and log_coefficients_mult

r^2 distribution
```{r}
cp_results_tidy |> 
  ggplot(aes(x = r.squared)) +
  geom_density()

```

r^2 distribution = partially left skewed. centers around 0.92. about 92% of the variation in tmax can be explained by the linear relationship between tmax and tmin + prcp

```{r}
cp_results_tidy |> 
  filter(log_coefficients_mult == "NaN") |> 
  nrow()

cp_results_tidy |> 
  filter(!log_coefficients_mult == "NaN") |> 
  ggplot(aes(x = log_coefficients_mult)) +
  geom_density()

```
noted total number of NAs produced; filtered them out for the distribution

log of product of coefficients distribution = centers around -5.5; What does this mean?

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2
 and log(β̂ 0∗β̂ 1)
. 
```{r}
cp_results_tidy |> 
  select(r.squared, log_coefficients_mult) |> 
  pivot_longer(
    r.squared:log_coefficients_mult,
    names_to = "term",
    values_to = "estimate"
  ) |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025, na.rm = TRUE),
    ci_upper = quantile(estimate, 0.975, na.rm = TRUE))
```


# Problem 3







